---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns

In SQL, you can filter columns by using SELECT and rows by using WHERE. Let's look at an example using the Lahman baseball database. First, we should load the Lahman package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example. Suppose we want to see the home run totals for the 1927 Yankees. We could write:

```{r}
query<-"SELECT playerID,teamID,yearID,HR from Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

We can also use the ">" and "<" keys to place parameters on the data we are given. For instance, if we only want to see instances of 1990 Phillies who hit more than 5 home runs, we can write:

```{r}
query<-"SELECT playerID,teamID,yearID,HR from Batting
WHERE teamID='PHI' and yearID=1990 and HR>5"
sqldf(query)
```

We can place as many conditions we want on the data output as long as we continue to place "and" before each condition. We can also use ">=" and "<=" to say "greater/less than or equal to".


## ORDER BY

With SQL, you can also order the rows in the output. For example, suppose you wanted to see every instance of a player hitting more than 50 home runs, but you would like the players with the most home runs to be at the top of this list. We could do this in the following way:

```{r}
query<-"SELECT playerID,teamID,yearID,HR from Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```
