---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns

In SQL, you can filter columns by using SELECT and rows by using WHERE. Let's look at an example using the Lahman baseball database. First, we should load the Lahman package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example. Suppose we want to see the home run totals for the 1927 Yankees. We could write:

```{r}
query<-"SELECT playerID,teamID,yearID,HR from Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

We can also use the ">" and "<" keys to place parameters on the data we are given. For instance, if we only want to see instances of 1990 Phillies who hit more than 5 home runs, we can write:

```{r}
query<-"SELECT playerID,teamID,yearID,HR from Batting
WHERE teamID='PHI' and yearID=1990 and HR>5"
sqldf(query)
```

We can place as many conditions we want on the data output as long as we continue to place "and" before each condition. We can also use ">=" and "<=" to say "greater/less than or equal to".


## ORDER BY

With SQL, you can also order the rows in the output. For example, suppose you wanted to see every instance of a player hitting more than 50 home runs, but you would like the players with the most home runs to be at the top of this list. We could do this in the following way:

```{r}
query<-"SELECT playerID,teamID,yearID,HR from Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```


## Aggregation

We can group our data and ask for sums of statistics that belong to the same group. For instance, if we want to see career sums of home runs, we can group the statistics by player ID, and then SQL will add each player's annual home run totals. Remember to use "HAVING" and not "WHERE" when placing conditions in this case.

```{r}
query<-"SELECT playerID,teamID,sum(HR)
FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY HR DESC"
sqldf(query)
```


## Joins

If you need information from multiple tables, you can construct a join. If we want to see records of Sammy Sosa's home runs throughout his career and want to see his name instead of an obscure player ID, we can join the professional statistics from the Batting table with the personal statistics of the Master table.

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='sosasa01'"
sqldf(query)
```